Equations & R Code

### 2017-02-18 manuscript/coexistence-stability_ms.Rmd
trying new model description

\renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}

\newcommand{\tc}{\textcolor}

\sloppy

\renewcommand{\baselinestretch}{1.25}
\usepackage[usenames]{color}
\newcommand{\new}{\textcolor{blue}}
<!--
\allsectionsfont{\normalfont\sffamily\bfseries}
-->

@@

MATERIALS AND METHODS
---------------------
###  Consumer-resource model
We developed a semi-discrete consumer-resource model that allows many species to coexist on one resource by either the storage effect or relative nonlinearity. 
In our model, the consumer can be in one of two-states: a dormant state $D$ and a live state $N$.
The dormant state could represent, for example, the seed bank of an annual plant.
Transitions between $N$ and $D$ occur at discrete intervals $\tau$ with continuous-time consumer-resource dynamics between discrete transitions. 
Thus, our model is formulated as "pulsed differential equations" [@Pachepsky2008; @Mailleret2009; @Mordecai2016]. 
For clarity we refer to $\tau$ as years and the growing time between years as seasons with daily ($t$) time steps.

During a growing season, consumer-resource dynamics are modeled as two differential equations:
\vspace{-3em}
\begin{align}
\frac{\text{d}N_{i}}{\text{d}t} &= \epsilon_if_{i}(R)N_{i}, \quad t \ne \tau_k\\
\frac{\text{d}R}{\text{d}t} &= - \sum\limits_{i}f_{i}(R)N_{i}, \quad t \ne \tau_k
\end{align}
\vspace{-3em}
\noindent where the discrete transitions between $N$ and $D$ occur between seasons at times $\tau_k$, $k = 1,2,3, \dots, K$. The subscript $i$ denotes species, $N$ is the living biomass state, and $\epsilon_i$ is each species' resource-to-biomass conversion efficiency. The growth rate of living biomass is a resource-dependent Hill function, $f_{i}(R) = r_{i}R^{a_{i}} / (b_{i}^{a_{i}}+R^{a_{i}})$, where *r* is a species' intrinsic growth rate and *a* and *b* define the curvature of the function. Resource depletion is equal to the sum of each species' consumption.

Along with resource uptake, consumer population growth depends on the production of dormant biomass ($D$), the activation of dormant biomass to live biomass ($D \rightarrow N$), and the survival of living biomass from one year to the next. The biomass of each species' states at the start of a growing season are equal to
\vspace{-3em}
\begin{align}
  D_{i}(\tau_k^+) &= (1-\gamma_{i,\tau_k})[\alpha_i N_{i}(\tau_k) + D_{i}(\tau_k)](1-\eta_i) \\
  N_{i}(\tau_k^+) &= (1-\alpha_i)N_{i}(\tau_k) + \gamma_{i,t}[\alpha_i N_{i}(\tau_k) + D_{i}(\tau_k)] (1-\eta_i),
\end{align}
\vspace{-3em}
\noindent where $D(\tau_k)$, $N(\tau_k)$, and $R(\tau_k)$ are the abundances of each state at the end of growing season *k* and $\tau_k^+$ denotes the beginning of growing season $k$. 
The activation of dormant biomass to live biomass is controlled by $\gamma$, which is year (*k*) and species (*i*) specific.
Dormant biomass is equal to a constant fraction ($\alpha$) of live biomass at the end of the previous season ($N_{i}(\tau_k)$), plus survival ($1-\eta_i$) of dormant biomass ($D_{i}(\tau_k)$) at the end of the previous year and dormant biomass remaining after live biomass activation ($D_{i}(\tau_k)(1-\gamma_{i,\tau_k})$).
Live biomass is equal to newly activated dormant biomass ($\gamma_{i,t}D_{i}(\tau_k)$), minus some fraction of live biomass that is converted to dormant biomass ($(1-\alpha_i)N_{i}(\tau_k)$)
We assume the resource pool is not replenished within a growing season.
Resource replenishment occurs between growing seasons, and the resource pool (*R*) at the start of the growing season *k*+1 is $R(\tau_k^+) = R^+$, where $R^+$ is a random resource pulse drawn from a log-normal distribution with mean $\mu(R^+)$ and standard deviation $\sigma(R^+)$. 
Model parameters and notation are described in Table 1.

>>>>

\renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}
\newcommand{\new}{\textcolor{blue}}
<!--
\allsectionsfont{\normalfont\sffamily\bfseries}
-->

@@

MATERIALS AND METHODS
---------------------
###  Consumer-resource model
We developed a semi-discrete consumer-resource model that allows many species to coexist on one resource by either the storage effect or relative nonlinearity. 
In our model, the consumer can be in one of two-states: a dormant state $D$ and a live state $N$.
The dormant state could represent, for example, the seed bank of an annual plant \new{or root biomass of a perennial plant}.
Transitions between $N$ and $D$ occur at discrete intervals $\tau$ with continuous-time consumer-resource dynamics between discrete transitions. 
Thus, our model is formulated as "pulsed differential equations" [@Pachepsky2008; @Mailleret2009; @Mordecai2016]. 
For clarity we refer to $\tau$ as years and the growing time between years as seasons with daily ($t$) time steps.

\new{At the beginning of growing season $\tau$ a season-specific fraction ($\gamma_{i,\tau}$) of dormant biomass is activated as living biomass such that}
\vspace{-3em}
\begin{align}
N_i(\tau^+) = \gamma_{i,\tau} D_i(\tau^+),
\end{align}
\vspace{-3em}
\noindent \new{where \emph{i} indexes each species and $\tau^+$ denotes the beginning of growing season $\tau$.}
\new{Live biomass at the start of the growing season ($N_i(\tau^+)$) then serves as the initial conditions for continuous-time} consumer-resource dynamics that are modeled as two differential equations:
\vspace{-3em}
\begin{align}
\frac{\text{d}N_{i}}{\text{d}t} &= \epsilon_if_{i}(R)N_{i}, \\
\frac{\text{d}R}{\text{d}t} &= - \sum\limits_{i}f_{i}(R)N_{i},
\end{align}
\vspace{-3em}
\noindent where the subscript $i$ denotes species, $N$ is the living biomass state, and $\epsilon_i$ is each species' resource-to-biomass conversion efficiency. The growth rate of living biomass is a resource-dependent Hill function, $f_{i}(R) = r_{i}R^{a_{i}} / (b_{i}^{a_{i}}+R^{a_{i}})$, where *r* is a species' intrinsic growth rate and *a* and *b* define the curvature of the function. Resource depletion is equal to the sum of each species' consumption.

\new{At the end of the growing season, a fraction ($\alpha_i$) of live biomass is stored as dormant biomass and a fraction of dormant biomass survives ($\eta_i$) to the next growing season, giving the following equation:}
\vspace{-3em}
\begin{align}
D_i{\tau + 1} = \left[\alpha_i N_i(\tau^-) + D_i(\tau^-) \right](1 - \eta_i)
\end{align}
\vspace{-3em}
\noindent \new{ where $\tau^-$ denotes the end of growing season $\tau$.
We assume remaining live biomass ($N(\tau^-)(1-\alpha)$) dies (i.e., this is not a closed system where all biomass must be in either $N$ or $D$ states).}

We assume the resource pool is not replenished within a growing season.
Resource replenishment occurs between growing seasons, and the resource pool (*R*) at the start of the growing season is $R(\tau^+) = R^+$, where $R^+$ is a random resource pulse drawn from a log-normal distribution with mean $\mu(R^+)$ and standard deviation $\sigma(R^+)$. 
\new{Taken all together, we can combine equations 1 and 4 to define the discrete transitions between live and dormant biomass at the end of a growing season.
Thus, the initial conditions for each state ($N, D, R$) at the beginning of growing season $\tau+1$ are:}
\vspace{-3em}
\begin{align}
  D_{i}(\tau^+ + 1) &= (1-\gamma_{i,\tau}) \left[\alpha_i N_{i}(\tau^-) + D_{i}(\tau^-) \right] (1-\eta_i) \\
  N_{i}(\tau^+ + 1) &= \gamma_{i,\tau} \left[\alpha_i N_{i}(\tau^-) + D_{i}(\tau^-) \right] (1-\eta_i) \\
  R(\tau^+ + 1) &= \text{lognormal}\left(\mu(R^+), \sigma(R^+) \right),
\end{align}
\vspace{-3em}
\noindent \new{where, as above, $\tau^-$ denotes the end of growing season $\tau$ and $\tau^+ + 1$ denotes the beginning of growing season $\tau+1$.} 
Model parameters and notation are described in Table 1.

### 2017-02-18 manuscript/coexistence-stability_ms.tex
trying new model description

\renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}

\newcommand{\tc}{\textcolor}

\sloppy

\renewcommand{\baselinestretch}{1.25}

@@

We developed a semi-discrete consumer-resource model that allows many
species to coexist on one resource by either the storage effect or
relative nonlinearity. In our model, the consumer can be in one of
two-states: a dormant state \(D\) and a live state \(N\). The dormant
state could represent, for example, the seed bank of an annual plant.
Transitions between \(N\) and \(D\) occur at discrete intervals \(\tau\)
with continuous-time consumer-resource dynamics between discrete
transitions. Thus, our model is formulated as ``pulsed differential
equations'' (Pachepsky et al. 2008; Mailleret \& Lemesle 2009; Mordecai
et al. 2016). For clarity we refer to \(\tau\) as years and the growing
time between years as seasons with daily (\(t\)) time steps.

During a growing season, consumer-resource dynamics are modeled as two
differential equations: \vspace{-3em}

\begin{align}
\frac{\text{d}N_{i}}{\text{d}t} &= \epsilon_if_{i}(R)N_{i}, \quad t \ne \tau_k\\
\frac{\text{d}R}{\text{d}t} &= - \sum\limits_{i}f_{i}(R)N_{i}, \quad t \ne \tau_k
\end{align}\vspace{-3em}

\noindent where the discrete transitions between \(N\) and \(D\) occur
between seasons at times \(\tau_k\), \(k = 1,2,3, \dots, K\). The
subscript \(i\) denotes species, \(N\) is the living biomass state, and
\(\epsilon_i\) is each species' resource-to-biomass conversion
efficiency. The growth rate of living biomass is a resource-dependent
Hill function,
\(f_{i}(R) = r_{i}R^{a_{i}} / (b_{i}^{a_{i}}+R^{a_{i}})\), where
\emph{r} is a species' intrinsic growth rate and \emph{a} and \emph{b}
define the curvature of the function. Resource depletion is equal to the
sum of each species' consumption.

Along with resource uptake, consumer population growth depends on the
production of dormant biomass (\(D\)), the activation of dormant biomass
to live biomass (\(D \rightarrow N\)), and the survival of living
biomass from one year to the next. The biomass of each species' states
at the start of a growing season are equal to \vspace{-3em}

\begin{align}
  D_{i}(\tau_k^+) &= (1-\gamma_{i,\tau_k})[\alpha_i N_{i}(\tau_k) + D_{i}(\tau_k)](1-\eta_i) \\
  N_{i}(\tau_k^+) &= (1-\alpha_i)N_{i}(\tau_k) + \gamma_{i,t}[\alpha_i N_{i}(\tau_k) + D_{i}(\tau_k)] (1-\eta_i),
\end{align}\vspace{-3em}

\noindent where \(D(\tau_k)\), \(N(\tau_k)\), and \(R(\tau_k)\) are the
abundances of each state at the end of growing season \emph{k} and
\(\tau_k^+\) denotes the beginning of growing season \(k\). The
activation of dormant biomass to live biomass is controlled by
\(\gamma\), which is year (\emph{k}) and species (\emph{i}) specific.
Dormant biomass is equal to a constant fraction (\(\alpha\)) of live
biomass at the end of the previous season (\(N_{i}(\tau_k)\)), plus
survival (\(1-\eta_i\)) of dormant biomass (\(D_{i}(\tau_k)\)) at the
end of the previous year and dormant biomass remaining after live
biomass activation (\(D_{i}(\tau_k)(1-\gamma_{i,\tau_k})\)). Live
biomass is equal to newly activated dormant biomass
(\(\gamma_{i,t}D_{i}(\tau_k)\)), minus some fraction of live biomass
that is converted to dormant biomass (\((1-\alpha_i)N_{i}(\tau_k)\)) We
assume the resource pool is not replenished within a growing season.
Resource replenishment occurs between growing seasons, and the resource
pool (\emph{R}) at the start of the growing season \emph{k}+1 is
\(R(\tau_k^+) = R^+\), where \(R^+\) is a random resource pulse drawn
from a log-normal distribution with mean \(\mu(R^+)\) and standard
deviation \(\sigma(R^+)\). Model parameters and notation are described
in Table 1.

>>>>

\renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}
\newcommand{\new}{\textcolor{blue}}

@@

We developed a semi-discrete consumer-resource model that allows many
species to coexist on one resource by either the storage effect or
relative nonlinearity. In our model, the consumer can be in one of
two-states: a dormant state \(D\) and a live state \(N\). The dormant
state could represent, for example, the seed bank of an annual plant
\new{or root biomass of a perennial plant}. Transitions between \(N\)
and \(D\) occur at discrete intervals \(\tau\) with continuous-time
consumer-resource dynamics between discrete transitions. Thus, our model
is formulated as ``pulsed differential equations'' (Pachepsky et al.
2008; Mailleret \& Lemesle 2009; Mordecai et al. 2016). For clarity we
refer to \(\tau\) as years and the growing time between years as seasons
with daily (\(t\)) time steps.

\new{At the beginning of growing season $\tau$ a season-specific fraction ($\gamma_{i,\tau}$) of dormant biomass is activated as living biomass such that}
\vspace{-3em}

\begin{align}
N_i(\tau^+) = \gamma_{i,\tau} D_i(\tau^+),
\end{align}\vspace{-3em}

\noindent \new{where \emph{i} indexes each species and $\tau^+$ denotes the beginning of growing season $\tau$.}
\new{Live biomass at the start of the growing season ($N_i(\tau^+)$) then serves as the initial conditions for continuous-time}
consumer-resource dynamics that are modeled as two differential
equations: \vspace{-3em}

\begin{align}
\frac{\text{d}N_{i}}{\text{d}t} &= \epsilon_if_{i}(R)N_{i}, \\
\frac{\text{d}R}{\text{d}t} &= - \sum\limits_{i}f_{i}(R)N_{i},
\end{align}\vspace{-3em}

\noindent where the subscript \(i\) denotes species, \(N\) is the living
biomass state, and \(\epsilon_i\) is each species' resource-to-biomass
conversion efficiency. The growth rate of living biomass is a
resource-dependent Hill function,
\(f_{i}(R) = r_{i}R^{a_{i}} / (b_{i}^{a_{i}}+R^{a_{i}})\), where
\emph{r} is a species' intrinsic growth rate and \emph{a} and \emph{b}
define the curvature of the function. Resource depletion is equal to the
sum of each species' consumption.

\new{At the end of the growing season, a fraction ($\alpha_i$) of live biomass is stored as dormant biomass and a fraction of dormant biomass survives ($\eta_i$) to the next growing season, giving the following equation:}
\vspace{-3em}

\begin{align}
D_i{\tau + 1} = \left[\alpha_i N_i(\tau^-) + D_i(\tau^-) \right](1 - \eta_i)
\end{align}\vspace{-3em}

\noindent \new{ where $\tau^-$ denotes the end of growing season $\tau$.
We assume remaining live biomass ($N(\tau^-)(1-\alpha)$) dies (i.e., this is not a closed system where all biomass must be in either $N$ or $D$ states).}

We assume the resource pool is not replenished within a growing season.
Resource replenishment occurs between growing seasons, and the resource
pool (\emph{R}) at the start of the growing season is
\(R(\tau^+) = R^+\), where \(R^+\) is a random resource pulse drawn from
a log-normal distribution with mean \(\mu(R^+)\) and standard deviation
\(\sigma(R^+)\).
\new{Taken all together, we can combine equations 1 and 4 to define the discrete transitions between live and dormant biomass at the end of a growing season.
Thus, the initial conditions for each state ($N, D, R$) at the beginning of growing season $\tau+1$ are:}
\vspace{-3em}

\begin{align}
  D_{i}(\tau^+ + 1) &= (1-\gamma_{i,\tau}) \left[\alpha_i N_{i}(\tau^-) + D_{i}(\tau^-) \right] (1-\eta_i) \\
  N_{i}(\tau^+ + 1) &= \gamma_{i,\tau} \left[\alpha_i N_{i}(\tau^-) + D_{i}(\tau^-) \right] (1-\eta_i) \\
  R(\tau^+ + 1) &= \text{lognormal}\left(\mu(R^+), \sigma(R^+) \right),
\end{align}\vspace{-3em}

\noindent \new{where, as above, $\tau^-$ denotes the end of growing season $\tau$ and $\tau^+ + 1$ denotes the beginning of growing season $\tau+1$.}
Model parameters and notation are described in Table 1.
